\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-11-18 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte

\begin{minipage} [c]  {5in} \scriptsize%
Notes: We use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level). Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.
\end{minipage}
\end{center}
\end{figure}

@@

As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -1.9\% for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
\cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture.
The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping. Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip

\paragraph{The share of additive costs: Time dynamics} As next step, we investigate whether additive costs are a long-lasting, i.e. a structural phenomenon. Figure \ref{fig:part_cout_additif} reports the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$.

\begin{figure}[htbp]
\caption{Share of additive costs $\beta$ (3-digit classification)}
\label{fig:part_cout_additif}
\begin{center}
 \includegraphics[width=12cm, height=7.5cm]{Figure1_share_of_additive_in_totalTC.jpg}
%{\footnotesize {OECD data} }
\begin{minipage} [c]  {5in} \scriptsize%
Note:The share of additive costs $\beta$ is reported on the $y-$axis. Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
\end{minipage}
\end{center}
\end{figure}

@@

In both transport modes, the additive component appears of sizable importance throughout the period.
Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
On the contrary, they represent a sizable and structural dimension of international transport costs.
If the share of additive costs in total costs roughly amounts to 44\% (30\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, with a different pattern across transport modes. In maritime transport, the share of additive costs fluctuates around the average, while it has been steeply declining since the end of the 1980s in air transport. This suggests that the transport costs decrease is more than proportionally due to additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.\smallskip

@@

\subsubsection{Robustness to the separability assumption}
In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, which considers (following \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $\tau_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approx. 200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.
For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows. Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
%\textbf{LP: Pb avec le nb d'observations. } GD : CELA A L’AIR BON ET ON LAISSE COMME CELA
%
%\begin{table}[htbp]
%  \centering
%  \caption{Robustness to the separability assumption (average over the period) - restricted sample}
%\begin{center}
%    \begin{tabular}{lcc|cc}
%    \hline \hline
%    \multicolumn{5}{c}{Mean value over 1974-2019} \\
% \multicolumn{5}{c}{3 digits} \\    \hline \hline
%   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
%   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
%    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
%in \%)}  \\
%    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
%    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
%    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
%    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
%    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
%\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
%    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
%    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
% \textit{Diagnostic test} & & &  &  \\
%    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
%    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
%    AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
%    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
%\# observations & 2381 & 2381 & 2798 & 2798 \\
%\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
%\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
%    \end{tabular}%
%\end{center}
%\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
%The additive term is expressed in fraction of fas price. 1989 omitted in 3-digit estimation for air.}}
%\label{tab:robustness_separability}
%\end{table}

>>>>

\begin{minipage} [c]  {5in} \scriptsize%
Notes: We use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level). Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$. 1989 trade by air is excluded.
\end{minipage}
\end{center}
\end{figure}

@@

As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -1.9\% for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
\cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture.
The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.}

%Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
%For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping. Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip

\paragraph{The share of additive costs: Time dynamics} Figure \ref{fig:part_cout_additif} reports the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$.

\begin{figure}[htbp]
\caption{Share of additive costs $\beta$ (baseline at 3-digit level)}
\label{fig:part_cout_additif}
\begin{center}
 \includegraphics[width=12cm, height=7.5cm]{Figure1_share_of_additive_in_totalTC.jpg}
%{\footnotesize {OECD data} }
\begin{minipage} [c]  {5in} \scriptsize%
Note: The dotted line is the quadratic trend.
\end{minipage}
\end{center}
\end{figure}

@@

In both transport modes, the additive component appears of sizable importance throughout the period.
Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs play a substantial role in total transport costs throughout the period. If the share of additive costs in total costs roughly amounts to 44\% (30\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, with a different pattern across transport modes. After a period of common increase in air and vessel from the mid 1970s to the end of the 1980s, it has steadily decreased for air transport. In contrast, the share of additive costs has continued to increase until the 2000s. Put differently, the total transport costs decrease is more than proportionally due to additive costs in air transport, while it is mostly driven by ad-valorem cost reduction in vessel transport.\smallskip

@@

\subsubsection{Robustness to the separability assumption}
Following \cite{Irrazabal_2015}, we have assumed that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the sector ($s(k)$) dimensions. In this section, we provide a robustness analysis to this separability assumption. That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $\tau_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis). This comes at the cost of having to estimate a much larger number of fixed effects (for approx. 200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation intractable in practice (all the more given our non-linear setting).
This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.
For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows. Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
%\textbf{LP: Pb avec le nb d'observations. } GD : CELA A L’AIR BON ET ON LAISSE COMME CELA
%
%\begin{table}[htbp]
%  \centering
%  \caption{Robustness to the separability assumption (average over the period) - restricted sample}
%\begin{center}
%    \begin{tabular}{lcc|cc}
%    \hline \hline
%    \multicolumn{5}{c}{Mean value over 1974-2019} \\
% \multicolumn{5}{c}{3 digits} \\    \hline \hline
%   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
%   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
%    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
%in \%)}  \\
%    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
%    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
%    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
%    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
%    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
%\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
%    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
%    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
% \textit{Diagnostic test} & & &  &  \\
%    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
%    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
%    AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
%    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
%\# observations & 2381 & 2381 & 2798 & 2798 \\
%\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
%\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
%    \end{tabular}%
%\end{center}
%\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
%The additive term is expressed in fraction of fas price. 1989 omitted in 3-digit estimation for air.}}
%\label{tab:robustness_separability}
%\end{table}
